### 2025-05-06 _tex/paper.tex
Built site for gh-pages

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrreprt}
\usepackage{xcolor}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games},
  pdfauthor={Bhavana Jonnalagadda; Riley Jones},
  pdfkeywords={social deduction games, persuasion modeling, Werewolf
Among Us dataset, large language models, multimodal analysis},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games}
\author{Bhavana Jonnalagadda \and Riley Jones}
\date{2025-05-06}
\begin{document}
\maketitle
\begin{abstract}
Abstract TODO
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}

\chapter{Introduction}\label{introduction}

\begin{itemize}
\item
  Social deduction games like \emph{Werewolf} offer a clear way to
  evaluate how agents deceive, persuade, and reason in group
  settings(\citeproc{ref-wikiwerewolf}{Wikipedia contributors 2024}). In
  these games, players have limited information, hidden identities, and
  must convince others while trying to figure out who is lying. These
  challenges closely match real life situations involving trust,
  negotiation, and manipulation.
\item
  We wanted to compare how humans and large language models (LLMs)
  handle these situations. To do this, we used a recent human dataset
  and constructed one from an LLM-based simulator:

  \begin{itemize}
  \tightlist
  \item
    \emph{Werewolf Among Us}
    (\citeproc{ref-laiWerewolfUsMultimodal2022}{Lai et al. 2022}), a
    collection of real human gameplay annotated with persuasion
    strategies,
  \item
    \emph{Werewolf Arena}
    (\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis, Friedhoff, and
    Chen 2024}), a simulated environment where LLM agents play the game
    autonomously.
  \item
    While both studies show Werewolf generates complex strategic
    language, neither compares human and LLM behavior directly.
  \end{itemize}
\item
  Our project addresses this gap. We analyzed transcripts from both
  datasets, matched them by role, round, and persuasion strategy, and
  compared how humans and LLMs lie, persuade, and detect deception.
\item
  By annotating utterances with the same set of persuasive strategies,
  we clearly show how synthetic agents differ from or resemble humans
  when navigating deception in adversarial group interactions.
\end{itemize}

\section{Related Work}\label{related-work}

\subsection{Multi-Agent LLMs}\label{multi-agent-llms}

\begin{itemize}
\tightlist
\item
  Among us game (\citeproc{ref-chiAMONGAGENTSEvaluatingLarge2024}{Chi,
  Mao, and Tang 2024})
\item
  Collective problem solving
  (\citeproc{ref-duLargeLanguageModels2024}{Du, Rajivan, and Gonzalez
  2024})

  \begin{itemize}
  \tightlist
  \item
    ``analyses indicate that LLM agent groups exhibit more
    disagreements, complex statements, and a propensity for positive
    statements compared to human groups''
  \end{itemize}
\item
  Govsim (\citeproc{ref-piattiCooperateCollapseEmergence2024}{Piatti et
  al. 2024})

  \begin{itemize}
  \tightlist
  \item
    ``In GOVSIM, a society of AI agents must collectively balance
    exploiting a common resource with sustaining it for future use. This
    environment enables the study of how ethical considerations,
    strategic planning, and negotiation skills impact cooperative
    outcomes.''
  \end{itemize}
\item
  All found similar themes

  \begin{itemize}
  \tightlist
  \item
    That LLMs are capable and good at understanding the rules
  \item
    That they can cooperate and be sneaky
  \end{itemize}
\end{itemize}

\subsection{LLMs and Werewolf}\label{llms-and-werewolf}

\begin{itemize}
\tightlist
\item
  Examination of improving werewolf by LLMs
  (\citeproc{ref-xuLanguageAgentsReinforcement2024}{Xu et al. 2024})

  \begin{itemize}
  \tightlist
  \item
    ``our agents use an LLM to perform deductive reasoning and generate
    a diverse set of action candidates. Then an RL policy trained to
    optimize the decision-making ability chooses an action from the
    candidates to play in the game. Extensive experiments show that our
    agents overcome the intrinsic bias and outperform existing LLM-based
    agents in the Werewolf game.''
  \end{itemize}
\item
  Werewolf Arena (\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis,
  Friedhoff, and Chen 2024})

  \begin{itemize}
  \tightlist
  \item
    Used in this paper
  \end{itemize}
\item
  Explicitly discuss how none of the exisiting LLM+Werewolf papers
  examine the differences/compare from a human dataset
\end{itemize}

\chapter{Methods}\label{methods}

\section{Data}\label{data}

\subsection{Werewolf Among Us Human
Dataset}\label{werewolf-among-us-human-dataset}

We used the \emph{Werewolf Among Us} dataset
(\citeproc{ref-laiWerewolfUsMultimodal2022}{Lai et al. 2022}), which
contains annotated dialogues from over 150 real games of One Night
Werewolf and Avalon. These games differ from classic Werewolf because
they have only one round of discussion and voting, do not eliminate
players during the game, and include many specialized roles beyond
Villager and Werewolf. The dataset provides detailed annotations of
persuasion strategies for each utterance, such as accusations, defenses,
and identity claims. We specifically used the textual transcriptions and
their strategy annotations for our comparisons.

\subsection{Werewolf Arena (LLM
Dataset)}\label{werewolf-arena-llm-dataset}

The \emph{Werewolf Arena} dataset
(\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis, Friedhoff, and Chen
2024}) features simulated classic Werewolf games played by LLM agents
without human intervention. Unlike the one-round human games, these
simulations involve multiple rounds alternating between night (secret
actions) and day (open discussion). Each LLM agent is assigned a role
(Villager, Werewolf, Seer, Doctor) and receives context-specific prompts
that guide their strategic interactions.

We ran simulations using five LLM models: GPT-4o, GPT-4.1, GPT-4o-mini,
DeepSeek-Chat, and DeepSeek-Reasoner. We experimented with two
configurations:

\begin{itemize}
\item
  8 players with 8 rounds of discussion
\item
  10 players with 6 rounds of discussion
\end{itemize}

We chose these configurations to give villagers more time to coordinate,
extending the dialogue to better observe persuasive behaviors. After
running the simulations, we manually annotated the LLM-generated
dialogue using the same persuasion categories as in the human dataset.

\section{Analysis}\label{analysis}

We standardized annotations across both datasets so they could be
directly compared. Our analyses focused on frequency distributions of
persuasion strategies, comparisons by role (villager versus werewolf),
and differences in how humans and LLMs applied these strategies.

\chapter{Results}\label{results}

\begin{figure}[H]

\centering{

\centering{

\begin{verbatim}
Unable to display output for mime type(s): text/html
\end{verbatim}

}

\subcaption{\label{fig-llm-winrounds-1}The LLM wins, by how many rounds
that partiticular game had.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-latex/Data-EDA_Comparison-fig-llm-winrounds-output-2.png}}

}

\subcaption{\label{fig-llm-winrounds-2}}

}

\caption{\label{fig-llm-winrounds}}

\end{figure}%

\textsubscript{Source:
\href{https://CUBoulder-DS.github.io/CSCI-5423-Final/Data/EDA_Comparison-preview.html\#cell-fig-llm-winrounds}{Werewolf
Among Us: Human vs LLM Analysis}}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-latex/Data-EDA_Comparison-fig-stratbyrole-hum-output-1.png}}

}

\caption{\label{fig-stratbyrole-hum}}

\end{figure}%

\textsubscript{Source:
\href{https://CUBoulder-DS.github.io/CSCI-5423-Final/Data/EDA_Comparison-preview.html\#cell-fig-stratbyrole-hum}{Werewolf
Among Us: Human vs LLM Analysis}}

\begin{figure}[H]

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{paper_files/figure-latex/Data-EDA_Comparison-fig-stratbyrole-llms-output-1.png}}

}

\caption{\label{fig-stratbyrole-llms}}

\end{figure}%

\textsubscript{Source:
\href{https://CUBoulder-DS.github.io/CSCI-5423-Final/Data/EDA_Comparison-preview.html\#cell-fig-stratbyrole-llms}{Werewolf
Among Us: Human vs LLM Analysis}}

\chapter{Discussion and Conclusion}\label{discussion-and-conclusion}

Interpret findings, discuss limitations, and propose future work.

\section{Limitations}\label{limitations}

\section{Future Work}\label{future-work}

\section{Summary}\label{summary}

Summarize contributions and insights from the project.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\chapter{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-bailisWerewolfArenaCase2024}
Bailis, Suma, Jane Friedhoff, and Feiyang Chen. 2024. {``Werewolf
{Arena}: {A Case Study} in {LLM Evaluation} via {Social Deduction}.''}
July 18, 2024. \url{https://doi.org/10.48550/arXiv.2407.13943}.

\bibitem[\citeproctext]{ref-chiAMONGAGENTSEvaluatingLarge2024}
Chi, Yizhou, Lingjun Mao, and Zineng Tang. 2024. {``{AMONGAGENTS}:
{Evaluating Large Language Models} in the {Interactive Text-Based Social
Deduction Game}.''} July 24, 2024.
\url{https://doi.org/10.48550/arXiv.2407.16521}.

\bibitem[\citeproctext]{ref-choRoundTableInvestigatingGroup2024}
Cho, Young-Min, Raphael Shu, Nilaksh Das, Tamer Alkhouli, Yi-An Lai,
Jason Cai, Monica Sunkara, and Yi Zhang. 2024. {``{RoundTable}:
{Investigating Group Decision-Making Mechanism} in {Multi-Agent
Collaboration}.''} November 11, 2024.
\url{https://doi.org/10.48550/arXiv.2411.07161}.

\bibitem[\citeproctext]{ref-duLargeLanguageModels2024}
Du, Yinuo, Prashanth Rajivan, and Cleotilde Gonzalez. 2024. {``Large
{Language Models} for {Collective Problem-Solving}: {Insights} into
{Group Consensus Decision-Making}.''} \emph{Proceedings of the Annual
Meeting of the Cognitive Science Society} 46 (0).
\url{https://escholarship.org/uc/item/6s060914}.

\bibitem[\citeproctext]{ref-laiWerewolfUsMultimodal2022}
Lai, Bolin, Hongxin Zhang, Miao Liu, Aryan Pariani, Fiona Ryan, Wenqi
Jia, Shirley Anugrah Hayati, James M. Rehg, and Diyi Yang. 2022.
{``Werewolf {Among Us}: {A Multimodal Dataset} for {Modeling Persuasion
Behaviors} in {Social Deduction Games}.''} December 16, 2022.
\url{https://doi.org/10.48550/arXiv.2212.08279}.

\bibitem[\citeproctext]{ref-piattiCooperateCollapseEmergence2024}
Piatti, Giorgio, Zhijing Jin, Max Kleiman-Weiner, Bernhard Schölkopf,
Mrinmaya Sachan, and Rada Mihalcea. 2024. {``Cooperate or {Collapse}:
{Emergence} of {Sustainable Cooperation} in a {Society} of {LLM
Agents}.''} \emph{Advances in Neural Information Processing Systems} 37
(December): 111715--59.
\url{https://proceedings.neurips.cc/paper_files/paper/2024/hash/ca9567d8ef6b2ea2da0d7eed57b933ee-Abstract-Conference.html}.

\bibitem[\citeproctext]{ref-wikiwerewolf}
Wikipedia contributors. 2024. {``Mafia (Party Game).''}
\url{https://en.wikipedia.org/wiki/Mafia_(party_game)}.

\bibitem[\citeproctext]{ref-xuLanguageAgentsReinforcement2024}
Xu, Zelai, Chao Yu, Fei Fang, Yu Wang, and Yi Wu. 2024. {``Language
{Agents} with {Reinforcement Learning} for {Strategic Play} in the
{Werewolf Game}.''} February 20, 2024.
\url{https://doi.org/10.48550/arXiv.2310.18940}.

\end{CSLReferences}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\chapter{Project Contributions}\label{project-contributions}

\textbf{Bhavana Jonnalagadda}:

\begin{itemize}
\tightlist
\item
  Paper framework (Quarto) setup
\item
  Github repo management
\item
  EDA on LLM dataset
\item
  Final comparison EDA and results analysis
\item
  Results section
\item
  Discussion and Conclusion section
\item
  Abstract
\end{itemize}

\textbf{Riley Jones}:

\begin{itemize}
\tightlist
\item
  EDA on human dataset
\item
  Werewolf Arena LLM simulation running and data aquisition
\item
  Introduction section
\item
  Methods section
\end{itemize}

\end{document}